---
title: "Week_1"
author: "Philipp Meisinger"
date: "`r Sys.Date()`"
output: html_document
---

### First weeks exercises:

Importiere den Datensatz "umsatzdaten_gekuerzt.csv", den Du über folgenden Link herunterladen kannst: <https://raw.githubusercontent.com/opencampus-sh/einfuehrung-in-data-science-und-ml/main/wetter.csv> Berechne die Gesamtdurchschnittstemperatur. Berechne die Durchschnittstemperatur für den Monat Juli. Vergleiche, ob die Monate Juli und Mai sich in ihrer Durchschnittstemperatur signifikant unterscheiden.

```{r load data}
library(tidyverse)
data <- read.csv("inc/umsatzdaten_gekuerzt.csv")
weather_data <- read.csv("inc/wetter.csv")

# weather_data is a data_frame, which contains vectors. Vectors can be accessed by referencing the '$' syntax.

```

## Total mean temperature of weather:

```{r}
# Referencing the column 'Temperatur' of struct 'weather_data' by using '$':
#  ´weather_data$Temperatur´
# using mean() function to get average of values in column 'Temperatur':
# The total average of temperature for the data set is: 
mean(weather_data$Temperatur)

```

## Total mean temperature of weather in July:

```{r}
#Adding column 'month' to data set: 
weather_data$Monat <- month(ymd(weather_data$Datum))

# Filtering out values for 'July':
weather_july <- dplyr::filter(weather_data, weather_data$Monat == 7)

# Accessing the 'Temperatur' columns and calculating the average over it:
mean_july <- mean(weather_july$Temperatur)

```

## Comparison between Mays' and Julys' temperatures, are they significantly different?
Firstly, let's discuss, what 'significant' means in this context. A significance test can be made to ensure, that the result is of significance and is unlikely to occur in natural circumstances.
We consider the entire data set we have for each month a sample of the globally existing data set of past or future values of temperature in these months. This way, we can determine our samplesize n.

``` {r}
# Firstly, let's select the month may just like we did with the month of July above:
weather_may <- dplyr::filter(weather_data, weather_data$Monat == 5)

# We can then calculate the average temperature of may:
mean_may <- mean(weather_may$Temperatur)

# Now we have both the average value of temperature of May and of July. Let's look at those values:
print("This is the average temperature during the month of May: ")
mean_may
print("This is the average temperature during the month of July: ")
mean_july

# Secondly, we need to figure out, how spread out the data is around the mean. An average value of 50 may be consisting of 2 values, the first = 0, the second = 100. 100 / 2 = 50. The standard deviation would tell us, that the spread is very big in this example. Let's figure out what the spread is like for our examples:

standard_deviation_may <- sd(weather_may$Temperatur, na.rm = TRUE)
standard_deviation_may

standard_deviation_july <- sd(weather_july$Temperatur, na.rm = TRUE)
standard_deviation_july

# The null-hypothesis here would be, that the both months have the same average temperatures. H_0 = May and July have the same average temperatures. H_1 = May and July have significantly different average temperatures. We will do this test by applying the t-test. In order to do that, we join both, so that we have a data set, which contains the Months (5 or 7) and lots of values for temperature corresponding to those month values.

weather_may <- dplyr::select(weather_may, Temperatur, Monat)
weather_july <- dplyr::select(weather_july, Temperatur, Monat)
new_data <- dplyr::full_join(weather_july, weather_may, by = join_by(Monat, Temperatur))

# If we want to be extremely accurate, we'd check our assumption, that the sample is 'normally distributed' (Normalverteilung):
# shapiro.test(new_data$Temperatur)

t.test(Temperatur ~ Monat, data = new_data)


## Links for research on this topic:
# https://www.youtube.com/watch?v=x1RFWHV2VUU
# https://studyflix.de/statistik/t-test-1584
# https://www.scribbr.de/statistik/standardabweichung/
# https://www.statology.org/standard-deviation-in-r/
# https://www.youtube.com/watch?v=p_NsoIXSC2s
# https://www.datanovia.com/en/lessons/t-test-in-r/


# Let's see how accurate that is concerning the Confidence intervals being negative

# Calculate Confidence Interval in R for Normal Distribution
# Confidence Interval Statistics
# 95 percent confidence interval so tails are .925

mean <- mean_may
stddev <- standard_deviation_may
n <- 248
error <- qnorm(0.975)*stddev/sqrt(n)
error
lower_bound <- mean - error
lower_bound
upper_bound <- mean + error
upper_bound

# Just for fun we are going to calculate the variance (Varianz) of both data sets.

variance_may <- standard_deviation_may^2
variance_may

variance_july <- standard_deviation_july^2
variance_july

### Now this doesn't seem to align with the Confidence intervals of the upper calculation. Instead, we are going to look at only one population, which is the data from July and then see, if the sample (mean_may) can be explained by that distribution.

t.test(weather_july$Temperatur, mu = mean_may, data = weather_july)

t.test(weather_may$Temperatur, mu = mean_july, data = weather_may)

## That looks more like it!! 
# We now can say, that the null hypothesis can be denied and we can assume (for now) that our hypothesis (H_1) stands true. There is a statistically significant difference in the average temperatures of may and july from this data set.

```
